An Analysis of the SPDY Protocol as a Candidate for HTTP/2.0

### 2012-12-19 paper/paper.tex
Lots more about spdy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SPDY}
\label{sec:background/spdy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The most interesting and likely to succeed proposal, SPDY is rather mature
at this point in time, having been implemented in both the most recent Chrome
and Firefox browsers in addition to having modules and patches for both the
Apache and Nginx web servers which allow them to speak SPDY.

@@

The most important parts of SPDY as of draft 3 (most recent complete version
\cite{spdy3}) are framing and streams.

\begin{description}
\item[Framing Layer] This binary protocol provides a layer on top of a reliable transport
protocol (recommended to be TCP) and provides a way for HTTP request/response
pairs to be issued and received out of order over a single transport layer
connection.  This provides an expected performance improvement of HTTP/1.1 by
not requiring multiple connections to the host to be made in order to
parallelize resource acquisition. There are 2 types of frames: control
frames and data frames. Control frames which are responsible for managing the
lifecycle of a stream whereas data frames simply transmit data.
\item[Streaming] Streams are bidirectional and are used to send out of order
responses to requests. Importantly, streams can have priorities set by the
creator of the stream.  This, in conjunction with server-push (defined next)
allows for more important assets (images, javascripts, stylesheets, etc.) to be
sent over the wires first.  An example of the usefulness of this feature is a
webpage with many images that are unimportant for the actual content and layout
of the page.  In that case, the images could be given a low priority and sent
last.
\item[Server-Push] Once a server receives a request for an asset, if it is aware
of closely connected assets, it may send them behind the original response,
without waiting for the client to request them.  This will pre-populate the
cache of the client and save a round trip time for each item.
\item[Header Compression with Dictionary] If there are many small assets on a
webpage, then a significant portion of the transmitted bytes will be due to HTTP
headers rather than the actual content.  If this can be compressed, much
bandwidth is saved~\cite{binoy}. Furthermore, bandwidth for large headers (such
as cookies) will be saved by assuming headers present for future requests on
the same stream and in new streams once initialized.
\end{description}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SPDY}
\label{sec:background/spdy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The most interesting and likely to succeed proposal, SPDY, is rather mature
at this point in time, having been implemented in both the most recent Chrome
and Firefox browsers in addition to having modules and patches for both the
Apache and Nginx web servers which allow them to speak SPDY.

@@

We will briefly outline the key features of SPDY.

\subsubsection{Framing Layer}
This binary protocol is a layer on top of an already reliable transport
protocol (recommended, but not required, to be SSL+TLS over TCP) and provides a
way for HTTP requests and responses to be issued and received out of order over
a single transport layer connection.  This provides an expected performance
improvement over HTTP/1.1 by allowing connections to warm up after experiencing
TCP's slow start. Longer connections will have more-accurate congestion window
sizes, and thus the performance of requests will both be fair for early
requests and fast for later requests.

The framing layer, while taking steps to improve the speed of connections,
maintains the semantics of HTTP.  For example, HTTP headers are sent in a
HEADERS frame, which is a control frame used to convey the headers that would
be sent over a normal HTTP connection.  There are several types of control
frames including ones to set-up and tear-down streams (see
\S\ref{sec:background/spdy/streaming}), to determine the latency of the
connection, and other logistics as well.  The remainder of the traffic is
conveyed as data frames.

\subsubsection{Streaming}
\label{sec:background/spdy/streaming}
Streams are either unidirectional or bidirectional channels that are used to
send multiplexed requests and responses.  Stream multiplexing differs from
HTTP/1.1's pipelining in one important way: HTTP/1.1 enforces a FIFO queue of
responses to requests whereas many SPDY streams' data may be interleaved.
SPDY's stream multiplexing suffices to eliminate the inherent head-of-line
blocking.

Additionally, streams are prioritized by the creator of the stream.  This, in
conjunction with server-push (defined next) allows for more important assets
(images, javascripts, stylesheets, etc.) to be sent over the wire first.  An
example of the usefulness of this feature is a webpage with many images that
are unimportant for the actual content and layout of the page.  In that case,
the images could be given a low priority and sent last.

\subsubsection{Server Push}
Once a server receives a request for an asset, if it is aware
of closely connected assets, it may send them behind the original response,
without waiting for the client to request them.  This will pre-populate the
cache of the client and save a round trip time for each item. Since server push
is an optimization for clients, it is up to the clients to make sure that it is
indeed making their browsing experience faster. For example, the SPDY
protocol\cite{spdy3} states that ``browsers MUST implement throttles to
protect against unreasonable push attacks.'' Or, browsers may deny to open a
server push stream to receive an asset if it is already cached.

\subsubsection{Header Compression with Dictionary}
If there are many small assets on a webpage, then with HTTP/1.1 a significant
portion of the transmitted bytes will be due to HTTP headers rather than the
actual content. One low-hanging fruit for optimization is to compress
commonly-seen ASCII headers using zlib, which will save lots of of overhead on
sites with lots of requests~\cite{binoy}.

The SPDY specification includes a pre-seeded zlib dictionary of common HTTP
headers, so that even the beginnings of connections can benefit from header
compression.

\subsubsection{Security}
Although the current SPDY specification does not require mandatory encryption
over SSL, it is a design goal of SPDY to improve security: The SPDY designers
``believe that the long-term future of the web depends on a secure network
connection''\cite{spdy-whitepaper}.
